---
Title: "Getting Started with DairyPredictR"
author: "Dr. Rakesh Nehra and Dr. Pallavi Choudhary"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with DairyPredictR}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction

**DairyPredictR** is an R package developed for machine learning-based prediction of dairy lactation milk yield. It provides an easy-to-use workflow for importing, preprocessing, training, evaluating, and comparing various models.

The package currently supports the following machine learning algorithms:

- Multiple Linear Regression (MLR)
- Random Forest (RF)
- Gradient Boosting Machine (GBM)
- Support Vector Machine (SVM)
- Artificial Neural Network (ANN)


# Installation

Install the package from the local source file.

```{r eval=FALSE}
install.packages("DairyPredictR_1.0.0.tar.gz",
                 repos = NULL,
                 type = "source")
```

Load the package.

```{r eval=FALSE}
library(DairyPredictR)
```

# Typical Workflow

The complete workflow consists of six simple steps.

```Raw Data
    │
    ▼
read_dairy()
    │
    ▼
clean_data()
    │
    ▼
split_data()
    │
    ▼
fit_model()
    │
    ▼
predict()
    │
    ▼
evaluate()```

# Step 1: Import Data

Read dairy production records.

```{r eval=FALSE}
milk <- read_dairy("milk.csv")
```

Supported formats include

- CSV
- XLS
- XLSX

# Step 2: Data Cleaning

Clean missing values and prepare the dataset.

```{r eval=FALSE}
milk <- clean_data(milk)
```

# Step 3: Split Data

Split the data into training and testing datasets.

```{r eval=FALSE}
split <- split_data(
  milk,
  train_size = 0.80,
  seed = 123
)
```

# Step 4: Train a Model

Train a Random Forest model.

```{r eval=FALSE}
rf_model <- fit_model(
  data = split$train,
  target = "MilkYield305",
  method = "rf"
)
```

Available methods are

- "linear"
- "rf"
- "gbm"
- "svm"
- "ann"

# Step 5: Generate Predictions

Predict milk yield for the testing dataset.

```{r eval=FALSE}
pred <- predict(
  rf_model,
  split$test
)
```

# Step 6: Evaluate Model Performance

Evaluate prediction accuracy.

```{r eval=FALSE}
evaluate(
  rf_model,
  split$test
)
```

Example output

| Metric | Value |
|---------|------:|
| R² | 0.78 |
| RMSE | 214.35 |
| MAE | 166.48 |
| MAPE | 7.12 |
| Bias | -1.45 |

# Compare Multiple Models

Compare several machine learning algorithms simultaneously.

```{r eval=FALSE}
results <- compare_models(
  split,
  target = "MilkYield305"
)
```

Example output

| Model | R² | RMSE | MAE |
|--------|---:|------:|------:|
| RF | 0.784 | 214.35 | 166.48 |
| SVM | 0.776 | 221.11 | 170.42 |
| GBM | 0.761 | 228.87 | 175.63 |
| Linear | 0.742 | 239.91 | 183.25 |
| ANN | 0.731 | 244.62 | 187.54 |

# Plot Performance

Generate publication-quality figures.

```{r eval=FALSE}
plot_performance(results)
```

Plots are produced for

- R²
- RMSE
- MAE
- MAPE
- Bias

Figures may also be exported in PNG, PDF, and TIFF formats.

# Package Workflow Summary

The recommended analysis workflow is

1. Import data
2. Clean data
3. Split dataset
4. Train machine learning model
5. Generate predictions
6. Evaluate performance
7. Compare algorithms
8. Create publication-quality figures

# Citation

If you use DairyPredictR in your research, please cite the package in your publication.

# License

DairyPredictR is distributed under the GPL-3 License.

# Author

**Dr. Rakesh Nehra**

Department of Animal Genetics and Breeding

Lala lajpat rai university of veterinary and animal sciences,

Hisar, Haryana, India

# Session Information

```{r}
sessionInfo()
```
